By Sample Summary Of RNA-Seq Data
========================================================

Male Rats From Michal Pravenec:
-----------------------

```{r,echo=FALSE}
rm(list=ls())
options(stringsAsFactors=FALSE)
setwd("~/Documents/NextGenSeq")

strains = read.table(file="sampleInfo/pravenecStrains.txt",sep="\t",header=FALSE)
tissues = c("brain","heart","liver","brown adipose")
numSamples = 4

malesPravenec = data.frame(strain=rep(strains$V1,each=4),sampleNum=rep(c(1:4),nrow(strains)))
malesPravenec$sample = paste(gsub("/Cub","",gsub("/Ola","",gsub("-","",malesPravenec$strain,fixed=TRUE),fixed=TRUE),fixed=TRUE),malesPravenec$sampleNum,sep="_")

malesPravenec$brain = "in freezer"
malesPravenec$heart = "in freezer"
malesPravenec$liver = "in freezer"
malesPravenec$bat = "in freezer"

##  Parental Heart Analysis - 4/14
malesPravenec$heart[malesPravenec$strain %in% c("BN-Lx/Cub","SHR/Ola")] = "total RNA sequenced (02Jun14); small RNA sequenced (19May14)"

##  Batch 1 RI Brains - 9/14
batch1 = paste(rep(c("HXB27","HXB7","HXB25","HXB17","HXB2","HXB13","BXH12","SHR"),each=2),rep(c(1:2),8),sep="_")
malesPravenec$brain[malesPravenec$sample %in% batch1] = "total RNA sequenced (05Sep14); small RNA sequenced (21Jul14)"
              
##  Batch 2 RI Brains - 9/14             
batch2 = c(paste(c("HXB1","HXB3","HXB4","HXB5","HXB10","HXB15","HXB18","HXB20","HXB21","HXB22","HXB23","HXB24","HXB29","HXB31","BXH2"),"_1",sep=""))
malesPravenec$brain[malesPravenec$sample %in% batch2] = "total RNA sequenced (24Sep14); small RNA sequenced (23Sep14)"
malesPravenec$brain[malesPravenec$sample=="SHR_1"] = "total RNA sequenced (05Sep14); small RNA sequenced (21Jul14); REPLICATE (same extraction/different library prep) - total RNA sequenced (24Sep14) and small RNA sequenced (23Sep14)"

##  Batch 3 RI Brains - 9/25
batch3 = c("BXH3_1","BXH3_2","BXH5_1","BXH6_1","BXH6_2","BXH8_1","BXH9_1","BXH9_2","BXH10_1","BXH10_2","BXH11_1","BXH11_2","BXH13_1","BNLx_1","BNLx_2")
malesPravenec$brain[malesPravenec$sample %in% batch3] = "total RNA sequenced (20Nov14)"
malesPravenec$brain[malesPravenec$sample=="SHR_1"] = paste(malesPravenec$brain[malesPravenec$sample=="SHR_1"],"; REPLICATE2 total RNA sequenced (20Nov14)",sep="")
```


```{r,results="asis",echo=FALSE}
#make table pretty for printing
forPrint = malesPravenec[,c("strain","sample","brain","heart","liver","bat")]
colnames(forPrint) = c("Strain","Sample ID","Brain","Left Ventricle","Liver","Brown Adipose Tissue")
kable(forPrint,align=rep("c",ncol(forPrint)))
```

### By Tissue Summary

```{r,echo=FALSE}
summary = c()
for(tissue in c("brain","liver","heart","bat")){
  numSamples = aggregate(malesPravenec[,tissue],by=list(malesPravenec$strain),function(a) sum(grepl("total RNA sequenced",a)))
  numSamples.small = aggregate(malesPravenec[,tissue],by=list(malesPravenec$strain),function(a) sum(grepl("small RNA sequenced",a)))
  summary = rbind(summary,data.frame(tissue=tissue,avg.long=mean(numSamples$x),atLeast1.long=sum(numSamples$x>0),all4.long = sum(numSamples$x==4),avg.small=mean(numSamples.small$x),atLeast1.small=sum(numSamples.small$x>0),all4.small = sum(numSamples.small$x==4)))
}

```

```{r,results="asis",echo=FALSE}
#make table pretty for printing
forPrint2 = summary
forPrint2$avg.long = sprintf("%0.2f",round(forPrint2$avg.long,2))
forPrint2$avg.small = sprintf("%0.2f",round(forPrint2$avg.small,2))
colnames(forPrint2) = c("Tissue","Average Number of Samples Per Strain (total RNA)","Number of Strains with At Least 1 Sample (total RNA)","Number of Strains with All 4 Samples Per Strain (total RNA)","Average Number of Samples Per Strain (small RNA)","Number of Strains with At Least 1 Sample (small RNA)","Number of Strains with All 4 Samples Per Strain (small RNA)")
kable(forPrint2,align=rep("c",ncol(forPrint2)))
```

### Batch Projections

```{r}
inbred = c("LE","F344","SR","LEW","WKY","BUF","SS","DA")
proj = malesPravenec[malesPravenec$sample!="HXB29_4",c("strain","sample")]
proj = rbind(proj,data.frame(strain=rep(inbred,each=4),sample=paste(rep(inbred,each=4),"_",c(1:4),sep="")))
proj = rbind(data.frame(proj,tissue="brain"),data.frame(proj,tissue="liver"))

########################
## Completed Batches  ##
########################

proj$batch1 = ((proj$sample %in% batch1) & proj$tissue=="brain")
proj$batch2 = ((proj$sample %in% c(batch2,"SHR_1")) & proj$tissue=="brain")
proj$batch3 = ((proj$sample %in% c(batch3,"SHR_1")) & proj$tissue=="brain")

#########################
##  In Progress Batch  ##
#########################

batch4 = c("BXH12_1","BXH12_2","HXB13_1","HXB13_2","HXB17_1","HXB17_2","HXB2_1","HXB2_2","HXB25_1","HXB25_2","HXB27_1","HXB27_2","HXB7_1","HXB7_2","SHR_1","SHR_2")

proj$batch4 = ((proj$sample %in% batch4) & proj$tissue=="liver")

#########################
##  Projected Batches  ##
#########################

#projected
priority = unique(c(batch2,batch3[!(batch3 %in% c("BXH3_2","BXH6_2","BXH9_2"))],"SHR_1","SHR_3","SHR_4","BNLx_3","BNLx_4"))
proj$batch5 = ((proj$sample %in% priority) & proj$tissue=="liver")

## Samples sequenced more than once (should only be SHR_1)
proj[rowSums(proj[,paste("batch",1:5,sep="")])>1,]

## Number of strains with at least one sample sequenced by tissue
byStrain = aggregate(rowSums(proj[,paste("batch",1:5,sep="")])>0,by=list(strain=proj$strain,tissue=proj$tissue),sum)
table(byStrain$tissue[byStrain$x>0])


#projected - batch6
notSeq = proj[rowSums(proj[,paste("batch",1:5,sep="")])==0,]
notSeq.brain = notSeq$sample[notSeq$tissue=="brain" & grepl("X",notSeq$strain)]

batch6 = c("SHR_1","SHR_3","SHR_4","BNLx_3","BNLx_4",notSeq.brain[grep("_2",notSeq.brain,fixed=TRUE)])
notSeq.brain = notSeq.brain[!(notSeq.brain %in% batch6)]
set.seed(28937)
batch6 = c(batch6,sample(notSeq.brain[grep("_3",notSeq.brain,fixed=TRUE)],size=(32-length(batch6))))

#projected - batch7
notSeq.liver = notSeq$sample[notSeq$tissue=="liver" & grepl("X",notSeq$strain)]
batch7 = c("SHR_1","BXH3_2","BXH6_2","BXH9_2",batch6[batch6 %in% notSeq.liver])
notSeq.liver = notSeq.liver[!(notSeq.liver %in% batch7)]
set.seed(59782)
batch7 = c(batch7,sample(notSeq.liver[grep("_3",notSeq.liver,fixed=TRUE)],size=(32-length(batch7))))

## update master matrix
proj$batch6 = ((proj$sample %in% batch6) & proj$tissue=="brain")
proj$batch7 = ((proj$sample %in% batch7) & proj$tissue=="liver")

## Samples sequenced more than once (should only be SHR_1)
proj[rowSums(proj[,paste("batch",1:7,sep="")])>1,]

## Number of strains with at least one sample sequenced by tissue
byStrain = aggregate(rowSums(proj[,paste("batch",1:7,sep="")])>0,by=list(strain=proj$strain,tissue=proj$tissue),sum)
table(byStrain$tissue[byStrain$x>1])
table(byStrain$tissue[byStrain$x>2])


#projected - batch8
notSeq = proj[rowSums(proj[,paste("batch",1:7,sep="")])==0,]
notSeq.brain = notSeq$sample[notSeq$tissue=="brain" & grepl("X",notSeq$strain)]

batch8 = c("SHR_1",notSeq.brain[grep("_3",notSeq.brain,fixed=TRUE)])
set.seed(289546)
batch8 = c(batch8,sample(notSeq.brain[grep("_4",notSeq.brain,fixed=TRUE)],size=(32-length(batch8))))

#projected - batch9 - liver
notSeq.liver = notSeq$sample[notSeq$tissue=="liver" & grepl("X",notSeq$strain)]
batch9 = c("SHR_1",batch8[batch8 %in% notSeq.liver])
notSeq.liver = notSeq.liver[!(notSeq.liver %in% batch9)]
set.seed(649897)
batch9 = c(batch9,sample(notSeq.liver,size=(32-length(batch9))))

## update master matrix
proj$batch8 = ((proj$sample %in% batch8) & proj$tissue=="brain")
proj$batch9 = ((proj$sample %in% batch9) & proj$tissue=="liver")

## Samples sequenced more than once (should only be SHR_1)
proj[rowSums(proj[,paste("batch",1:9,sep="")])>1,]

## Number of strains with at least one sample sequenced by tissue
byStrain = aggregate(rowSums(proj[,paste("batch",1:9,sep="")])>0,by=list(strain=proj$strain,tissue=proj$tissue),sum)
table(byStrain$tissue[byStrain$x>2])
table(byStrain$tissue[byStrain$x>3])



#projected - batch10
notSeq = proj[rowSums(proj[,paste("batch",1:9,sep="")])==0,]
notSeq.brain = notSeq$sample[notSeq$tissue=="brain" & grepl("X",notSeq$strain)]
notSeq.inbred = notSeq$sample[notSeq$tissue=="brain" & !grepl("X",notSeq$strain)]

batch10 = c("SHR_1",notSeq.brain,notSeq.inbred[grep("_1",notSeq.inbred,fixed=TRUE)])
set.seed(1785)
batch10 = c(batch10,sample(notSeq.inbred[grep("_2",notSeq.inbred,fixed=TRUE)],size=(32-length(batch10))))

#projected - batch9 - liver
notSeq.liver = notSeq$sample[notSeq$tissue=="liver"]
batch11 = c("SHR_1",batch10[batch10 %in% notSeq.liver])
notSeq.liver = notSeq.liver[!(notSeq.liver %in% batch11)]
batch11 = c(batch11,sample(notSeq.liver[grep("_2",notSeq.liver,fixed=TRUE)],size=(32-length(batch11))))



## update master matrix
proj$batch10 = ((proj$sample %in% batch10) & proj$tissue=="brain")
proj$batch11 = ((proj$sample %in% batch11) & proj$tissue=="liver")

## Samples sequenced more than once (should only be SHR_1)
proj[rowSums(proj[,paste("batch",1:11,sep="")])>1,]

## Number of strains with at least one sample sequenced by tissue
byStrain = aggregate(rowSums(proj[,paste("batch",1:11,sep="")])>0,by=list(strain=proj$strain,tissue=proj$tissue),sum)
table(byStrain$tissue[byStrain$x>0])
table(byStrain$tissue[byStrain$x>1])
table(byStrain$tissue[byStrain$x>3])


#projected - batch12
notSeq = proj[rowSums(proj[,paste("batch",1:11,sep="")])==0,]


notSeq.brain = notSeq$sample[notSeq$tissue=="brain" & grepl("X",notSeq$strain)]








