---
title: "Comparison of smRNA yields"
author: "Laura Saba"
date: "July 1, 2015"
output: pdf_document
---


```{r}
rm(list=ls())
options(stringsAsFactors=FALSE)
library(knitr)

base.folder="~/Documents/NextGenSeq/"

tissue = "Liver"
batch = 2
file = "150602_7001413_0268_BC6K12ANXX_L4678_Tabakoff_demux.csv"

importQC = function(tissue,batch,file,date){
  x = read.table(file=paste(base.folder,"Rat.RI.",tissue,".smallRNA.batch",batch,"/data/",file,sep=""),skip=5,sep=",")
  colnames(x) = c("Lane","SampleID","SampleRef","Index","Description","Control","Project","Yield.Mbases","pctPF","numReads","pctRawClustPerLane","pctPerfectIndexReads","pctOneMismatchIndexReads","pctGrtQ30","MeanQualityScore")
  x$tissue = tissue
  x$batch = batch
  x$date = date
  return(x)
}

liverQC = rbind(importQC(tissue="Liver",batch=1,file="150219_7001413_0222_AC6G34ANXX_L1234_Tabakoff_demux.csv",date='25Feb15'),importQC(tissue="Liver",batch=2,file="150602_7001413_0268_BC6K12ANXX_L4678_Tabakoff_demux.csv",date='18Jun15'))

brainQC=rbind(importQC(tissue="Brain",batch=1,file="140702_7001413_0149_BC54PHACXX_L5678_Tabakoff_demux.csv",date='21Jul14'),importQC(tissue="Brain",batch=2,file="140918_SN634_0254_AC56FUACXX_L4567_Tabakoff_demux.csv",date='24Sep14'),importQC(tissue="Brain",batch=3,file="150107_7001413_0207_AC66N3ANXX_L2345_Tabakoff_demux.csv",date='16Jan15'),importQC(tissue="Brain",batch=4,file="150508_7001413_0258_AC6K0DANXX_L2_Tabakoff_demux.csv",date='14May15'))

totalQC = rbind(liverQC,brainQC)
totalQC = totalQC[totalQC$SampleRef!="unknown",]
totalQC$reads = as.numeric(gsub(",","",totalQC$numReads,fixed=TRUE))

boxplot(as.numeric(gsub(",","",totalQC$numReads,fixed=TRUE)) ~ totalQC$tissue*totalQC$batch)
```

By Batch Summary
```{r}
batchSummary = data.frame(aggregate(totalQC$reads,by=list(date=totalQC$date,batch=totalQC$batch,tissue=totalQC$tissue),min),numSamples =aggregate(totalQC$reads,by=list(date=totalQC$date,batch=totalQC$batch,tissue=totalQC$tissue),length)$x,numUnusable=aggregate(totalQC$reads,by=list(date=totalQC$date,batch=totalQC$batch,tissue=totalQC$tissue),function(a) sum(a<10000000))$x)

batchSummary$date = as.Date(batchSummary$date,"%d%b%y")

batchSummary[order(batchSummary$date),]
```

By Lane Summary
```{r}
laneSummary = data.frame(aggregate(totalQC$reads,by=list(lane=totalQC$Lane,date=totalQC$date,batch=totalQC$batch,tissue=totalQC$tissue),min),totalReads=prettyNum(aggregate(totalQC$reads,by=list(lane=totalQC$Lane,date=totalQC$date,batch=totalQC$batch,tissue=totalQC$tissue),sum)$x,big.mark=","))

laneSummary$date = as.Date(laneSummary$date,"%d%b%y")
laneSummary[order(laneSummary$date),]


