PANTHER RESULTS - 02/27/15
==========================
### Background
Mapped IDs:   14865  
Unmapped IDs: 	1089  

### Pink Module  
Mapped IDs: 	320  
Unmapped IDs: 	24  

```{r,echo=FALSE}
rm(list=ls())
options(stringsAsFactors=FALSE)
#library(WGCNA)

#base.path = "/data2/saba/BxH.HxB.Rats/RNA-Seq/"
base.path = "/Volumes/saba/BxH.HxB.Rats/RNA-Seq/"

goMF = read.table(file=paste(base.path,"data/PANTHER.Pink/goMolecularFunction.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goMF = goMF[goMF[,3]>1,]
goMF$pvalue = as.numeric(goMF[,5]=="+")*goMF[,7] + as.numeric(goMF[,5]=="-")*(1-goMF[,7])
goMF$fdr = p.adjust(goMF$pvalue,method="BH")

goBP = read.table(file=paste(base.path,"data/PANTHER.Pink/goBiologicalProcesses.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goBP = goBP[goBP[,3]>1,]
goBP$pvalue = as.numeric(goBP[,5]=="+")*goBP[,7] + as.numeric(goBP[,5]=="-")*(1-goBP[,7])
goBP$fdr = p.adjust(goBP$pvalue,method="BH")

goCC = read.table(file=paste(base.path,"data/PANTHER.Pink/goCellularComponent.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goCC = goCC[goCC[,3]>1,]
goCC$pvalue = as.numeric(goCC[,5]=="+")*goCC[,7] + as.numeric(goCC[,5]=="-")*(1-goCC[,7])
goCC$fdr = p.adjust(goCC$pvalue,method="BH")

colnames(goMF) = colnames(goBP) = colnames(goCC) = c("GO","bkgrd.Num","pink.Num","expected","direction","foldEnrich","two.sided.p","pvalue","fdr")

printMF = goMF[goMF$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]
printBP = goBP[goBP$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]
printCC = goCC[goCC$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]

printMF$fdr = sprintf("%.3f",round(printMF$fdr,3))
printBP$fdr = sprintf("%.3f",round(printBP$fdr,3))
printCC$fdr = sprintf("%.3f",round(printCC$fdr,3))

printMF$pvalue = sprintf("%.4f",round(printMF$pvalue,4))
printBP$pvalue = sprintf("%.4f",round(printBP$pvalue,4))
printCC$pvalue = sprintf("%.4f",round(printCC$pvalue,4))
```

### Enriched GO - Molecular Functions in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printMF,align=rep("c",ncol(printMF)))
```

### Enriched GO - Biological Processes in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printBP,align=rep("c",ncol(printBP)))
```

### Enriched GO - Cellular Components in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printCC,align=rep("c",ncol(printCC)))
```


```{r,echo=FALSE}
base.path = "/Volumes/saba/BxH.HxB.Rats/RNA-Seq/"

paths = read.table(file=paste(base.path,"data/PANTHER.Pink/pantherPathway.pink.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
paths = paths[paths[,3]>1,]
paths$pvalue = as.numeric(paths[,5]=="+")*paths[,7] + as.numeric(paths[,5]=="-")*(1-paths[,7])
paths$fdr = p.adjust(paths$pvalue,method="BH")

colnames(paths) = c("PANTHER Category","bkgrd.Num","pink.Num","expected","direction","foldEnrich","two.sided.p","pvalue","fdr")

printPaths = paths[paths$fdr<0.10,c("PANTHER Category","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]

printPaths$fdr = sprintf("%.3f",round(printPaths$fdr,3))
printPaths$pvalue = sprintf("%.4f",round(printPaths$pvalue,4))
```

### Enriched PANTHER Pathways in Pink Module with Nos1 (fdr<0.10)
```{r, results='asis',echo=FALSE}
kable(printPaths,align=rep("c",ncol(printPaths)))
```









