BNLx and SHR (Brain) small RNA RNA-Seq
========================================================

* 3 samples per strain 
* tissues received from ?
* small RNA (<200 bp)
* 1X50 reads using the unstranded protocol
* sequencing done by ?

1. Determine number of reads sent for each sample 
-------------------------------------------------
```
qsub -q compute -N brain.raw /storage/saba/BNLx.SHR.Brain.smallRNA/programs/rawReadCounts.brain.smallRNA.sh
```

```{r,echo=FALSE}
rm(list=ls())
options(stringsAsFactors=FALSE)
setwd("~/Documents/NextGenSeq/BNLx.SHR.Brain.smallRNA/")
rawCounts = read.table(file="data/rawReadCounts.03Mar15.txt",sep="\t",header=FALSE,fill=TRUE)
#rawCounts = rawCounts[!grepl("total",rawCounts$V1),]

rawCounts$reads = unlist(lapply(strsplit(sub("^\\s+", "",rawCounts[,1]),split=" ",fixed=TRUE),function(x) as.numeric(x[1])/4))
rawCounts$file = unlist(lapply(strsplit(sub("^\\s+", "",rawCounts[,1]),split=" ",fixed=TRUE),function(x) x[2]))
rawCounts$sample = unlist(lapply(strsplit(rawCounts$file,split="/",fixed=TRUE),function(x) gsub("small","",gsub(".fastq","",x[length(x)]))))
rawCounts = rawCounts[!grepl("total",rawCounts$sample),]

forPrint = rawCounts[,c("sample","reads")]
forPrint$reads = prettyNum(forPrint$reads,big.mark=",",scientific=FALSE)
colnames(forPrint) = c("sample","Number of Reads")
```

Raw Reads/Read Fragments
---------------------------

```{r, results='asis',echo=FALSE}
kable(forPrint,align=rep("c",ncol(forPrint)),row.names=FALSE)
```

Total Number of Reads:  `r prettyNum(sum(rawCounts$reads),big.mark=",",scientific=FALSE)`  
Average Number of Paired End Reads Per Sample: `r prettyNum(sum(rawCounts$reads)/nrow(rawCounts),big.mark=",",scientific=FALSE)`  



