Examination of WGCNA Module with Nitric Oxide Synthase 1
========================================================

```{r,echo=FALSE,eval=FALSE}
rm(list=ls())
options(stringsAsFactors=FALSE)
library(WGCNA)

#base.path = "/data2/saba/BxH.HxB.Rats/RNA-Seq/"
base.path = "/Volumes/saba/BxH.HxB.Rats/RNA-Seq/"

load(paste(base.path,"Rdata/strainEst.RUVs2.Rdata",sep=""))

dim(strainEst)

## limit to ensembl genes and recombinant inbred strains
datExpr = t(strainEst[grep("ENS",rownames(strainEst)),grepl("BXH",colnames(strainEst)) | grepl("HXB",colnames(strainEst))])

## determine appropriate beta parameter
powers = c(c(1:15),seq(from=16,to=20,by=2))

#sft=pickSoftThreshold(datExpr,powerVector=powers,corFnc="cor",corOptions=list(use ='pairwise.complete.obs', method = 'spearman'),verbose=5)
sft=pickSoftThreshold(datExpr,powerVector=powers,corFnc="bicor",verbose=5)

sizeGrWindow(9, 5)
par(mfrow = c(1,2));
cex1 = 0.8;


pdf(file=paste(base.path,"images/SoftThresholdingPower.24Feb15.pdf",sep=""))
plot(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
xlab="Soft Threshold (power)",ylab="Scale Free Topology Model Fit,signed R^2",type="n",
main = paste("Scale independence"));
text(sft$fitIndices[,1], -sign(sft$fitIndices[,3])*sft$fitIndices[,2],
labels=powers,cex=cex1,col="red")
# this line corresponds to using an R^2 cut-off of h
abline(h=0.80,col="red")

# Mean connectivity as a function of the soft-thresholding power
plot(sft$fitIndices[,1], sft$fitIndices[,5],
xlab="Soft Threshold (power)",ylab="Mean Connectivity", type="n",
main = paste("Mean connectivity"))
text(sft$fitIndices[,1], sft$fitIndices[,5], labels=powers, cex=cex1,col="red")

dev.off()

save(sft, file=paste(base.path,"Rdata/SoftThresholdingPower.23Feb15.Rdata",sep=""))


##############################################################
##  WGCNA in HXB Brain Expression Data (RNAseq Derived TC)  ##
##############################################################

net = blockwiseModules(datExpr,power=9, minModuleSize=5, deepSplit=4, numericLabels=TRUE,pamRespectsDendro=FALSE,saveTOMs=TRUE, saveTOMFileBase="brain",verbose=3,networkType="unsigned", corType="bicor")

moduleLabels = net$colors
moduleColors = labels2colors(net$colors)
names(moduleColors) = colnames(datExpr)
TC = colnames(datExpr)

moduleMembership = cbind(TC, moduleLabels, moduleColors)
MEs = net$MEs

save(net,moduleLabels,moduleColors,MEs,moduleMembership,file=paste(base.path,"Rdata/HXBbrain.RNASeq.WGCNA.27Feb15.Rdata",sep=""))


### Nos1 - ENSRNOG00000001130 ###

nos1.module = moduleMembership[moduleMembership[,1]=="ENSRNOG00000001130","moduleColors"]

## number of genes in module 
sum(moduleColors==nos1.module)
# 351 genes

nos1.corrs = data.frame(corNos1=bicor(datExpr,datExpr[,"ENSRNOG00000001130"]))
nos1.corrs$adjNos1 = abs(nos1.corrs$corNos1^9)

anno=read.table(file="/Volumes/saba/annotation/ensemblGenes.Rnor_5.0.08Jul14.txt",sep="\t",header=TRUE)
anno=anno[!duplicated(anno$EnsemblGeneID),]

nos1.corrs=merge(nos1.corrs,anno,by.x=0,by.y="EnsemblGeneID",all.x=TRUE)
nos1.corrs$adjNos1Rank = nrow(nos1.corrs)-rank(nos1.corrs$adjNos1)+1

nos1.corrs = merge(nos1.corrs,data.frame(module=moduleColors),by.x="Row.names",by.y=0)

nearestNeighbors = nos1.corrs[nos1.corrs$module=="pink" & nos1.corrs$GeneSymbol!="Nos1",]
nearestNeighbors = nearestNeighbors[order(nearestNeighbors$adjNos1Rank),]
topTen = nearestNeighbors[1:10,c("GeneSymbol","corNos1","Description")]

topTen$corNos1 = sprintf("%.2f",round(topTen$corNos1,2))
topTen$Description = unlist(lapply(strsplit(topTen$Description,split="[",fixed=TRUE),function(a) a[1]))
topTen$Description = unlist(lapply(strsplit(topTen$Description,split="(",fixed=TRUE),function(a) a[1]))


###  Examine Pink Module  ###
intraConnect = intramodularConnectivity.fromExpr(datExpr, moduleColors, corFnc = "bicor", networkType = "unsigned", power = 9)

intraConnect = data.frame(gene=colnames(datExpr),module=moduleColors,intraConnect)

pink.intra = intraConnect[intraConnect$module=="pink",]
pink.intra = merge(pink.intra,anno,by.x="gene",by.y="EnsemblGeneID",all.x=TRUE)
pink.intra$rank = nrow(pink.intra) - rank(pink.intra$kWithin) + 1

head(pink.intra[order(pink.intra$rank),c("gene","GeneSymbol","Description","kWithin","rank")],10)

pink.intra[pink.intra$GeneSymbol=="Nos1",]

write.table(unique(pink.intra$GeneSymbol[pink.intra$GeneSymbol!=""]),file=paste(base.path,"data/pinkModuleGeneSymbols.txt",sep=""),sep="\t",row.names=FALSE,col.names=FALSE,quote=FALSE)
write.table(unique(nos1.corrs$GeneSymbol[nos1.corrs$GeneSymbol!=""]),file=paste(base.path,"data/backgrdGeneSymbols.txt",sep=""),sep="\t",row.names=FALSE,col.names=FALSE,quote=FALSE)

```


Eigengene Characteristics
```{r}
head(MEs)

moduleMembership[moduleMembership[,1]=="ENSRNOG00000001130",]

pinkME = MEs[,"ME8"]
names(pinkME) = rownames(datExpr)
pinkME = pinkME[order(pinkME)]  

pdf(file=paste(base.path,"images/eigengeneDist.27Feb15.pdf",sep=""))
eigenPlot = barplot(pinkME,ylab="Pink Eigengene Value",xaxt="n", xlab="")  
axis(1,at=eigenPlot,labels=FALSE)
text(eigenPlot, par("usr")[3] - 0.08, labels = names(pinkME), srt = 90, xpd = TRUE,cex=0.7)
dev.off()
  
PANTHER RESULTS - 02/27/15
==========================
### Background
Mapped IDs: 	14865  
Unmapped IDs: 	1089  

### Pink Module  
Mapped IDs: 	320  
Unmapped IDs: 	24  

```{r,echo=FALSE}
rm(list=ls())
options(stringsAsFactors=FALSE)
#library(WGCNA)

#base.path = "/data2/saba/BxH.HxB.Rats/RNA-Seq/"
base.path = "/Volumes/saba/BxH.HxB.Rats/RNA-Seq/"

goMF = read.table(file=paste(base.path,"data/PANTHER.Pink/goMolecularFunction.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goMF = goMF[goMF[,3]>1,]
goMF$pvalue = as.numeric(goMF[,5]=="+")*goMF[,7] + as.numeric(goMF[,5]=="-")*(1-goMF[,7])
goMF$fdr = p.adjust(goMF$pvalue,method="BH")

goBP = read.table(file=paste(base.path,"data/PANTHER.Pink/goBiologicalProcesses.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goBP = goBP[goBP[,3]>1,]
goBP$pvalue = as.numeric(goBP[,5]=="+")*goBP[,7] + as.numeric(goBP[,5]=="-")*(1-goBP[,7])
goBP$fdr = p.adjust(goBP$pvalue,method="BH")

goCC = read.table(file=paste(base.path,"data/PANTHER.Pink/goCellularComponent.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goCC = goCC[goCC[,3]>1,]
goCC$pvalue = as.numeric(goCC[,5]=="+")*goCC[,7] + as.numeric(goCC[,5]=="-")*(1-goCC[,7])
goCC$fdr = p.adjust(goCC$pvalue,method="BH")

colnames(goMF) = colnames(goBP) = colnames(goCC) = c("GO","bkgrd.Num","pink.Num","expected","direction","foldEnrich","two.sided.p","pvalue","fdr")

printMF = goMF[goMF$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]
printBP = goBP[goBP$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]
printCC = goCC[goCC$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]

printMF$fdr = sprintf("%.3f",round(printMF$fdr,3))
printBP$fdr = sprintf("%.3f",round(printBP$fdr,3))
printCC$fdr = sprintf("%.3f",round(printCC$fdr,3))

printMF$pvalue = sprintf("%.4f",round(printMF$pvalue,4))
printBP$pvalue = sprintf("%.4f",round(printBP$pvalue,4))
printCC$pvalue = sprintf("%.4f",round(printCC$pvalue,4))
```

### Enriched GO - Molecular Functions in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printMF,align=rep("c",ncol(printMF)))
```

### Enriched GO - Biological Processes in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printBP,align=rep("c",ncol(printBP)))
```

### Enriched GO - Cellular Components in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printCC,align=rep("c",ncol(printCC)))
```

```{r,echo=FALSE}
base.path = "/Volumes/saba/BxH.HxB.Rats/RNA-Seq/"

goMFExp = read.table(file=paste(base.path,"data/PANTHER.Pink/goMolecularFunctionExperimental.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goMFExp = goMFExp[goMFExp[,3]>1,]
goMFExp$pvalue = as.numeric(goMFExp[,5]=="+")*goMFExp[,7] + as.numeric(goMFExp[,5]=="-")*(1-goMFExp[,7])
goMFExp$fdr = p.adjust(goMFExp$pvalue,method="BH")

goBPExp = read.table(file=paste(base.path,"data/PANTHER.Pink/goBiologicalProcessesExperimental.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goBPExp = goBPExp[goBPExp[,3]>1,]
goBPExp$pvalue = as.numeric(goBPExp[,5]=="+")*goBPExp[,7] + as.numeric(goBPExp[,5]=="-")*(1-goBPExp[,7])
goBPExp$fdr = p.adjust(goBPExp$pvalue,method="BH")

goCCExp = read.table(file=paste(base.path,"data/PANTHER.Pink/goCellularComponentExperimental.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
goCCExp = goCCExp[goCCExp[,3]>1,]
goCCExp$pvalue = as.numeric(goCCExp[,5]=="+")*goCCExp[,7] + as.numeric(goCCExp[,5]=="-")*(1-goCCExp[,7])
goCCExp$fdr = p.adjust(goCCExp$pvalue,method="BH")

colnames(goMFExp) = colnames(goBPExp) = colnames(goCCExp) = c("GO","bkgrd.Num","pink.Num","expected","direction","foldEnrich","two.sided.p","pvalue","fdr")

printMFExp = goMFExp[goMFExp$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]
printBPExp = goBPExp[goBPExp$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]
printCCExp = goCCExp[goCCExp$fdr<0.10,c("GO","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]

printMFExp$fdr = sprintf("%.3f",round(printMFExp$fdr,3))
printBPExp$fdr = sprintf("%.3f",round(printBPExp$fdr,3))
printCCExp$fdr = sprintf("%.3f",round(printCCExp$fdr,3))

printMFExp$pvalue = sprintf("%.4f",round(printMFExp$pvalue,4))
printBPExp$pvalue = sprintf("%.4f",round(printBPExp$pvalue,4))
printCCExp$pvalue = sprintf("%.4f",round(printCCExp$pvalue,4))
```

### Enriched GO - Molecular Functions (Experimental) in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printMFExp,align=rep("c",ncol(printMFExp)))
```

### Enriched GO - Biological Processes (Experimental) in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printBPExp,align=rep("c",ncol(printBPExp)))
```

### Enriched GO - Cellular Components (Experimental) in Pink Module with Nos1 (FDR<0.10)
```{r, results='asis',echo=FALSE}
kable(printCCExp,align=rep("c",ncol(printCCExp)))
```


```{r,echo=FALSE}
base.path = "/Volumes/saba/BxH.HxB.Rats/RNA-Seq/"

paths = read.table(file=paste(base.path,"data/PANTHER.Pink/pantherPathway.pink.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
paths = paths[paths[,3]>1,]
paths$pvalue = as.numeric(paths[,5]=="+")*paths[,7] + as.numeric(paths[,5]=="-")*(1-paths[,7])
paths$fdr = p.adjust(paths$pvalue,method="BH")

proteinClass = read.table(file=paste(base.path,"data/PANTHER.Pink/pantherProteinClass.27Feb15.txt",sep=""),sep="\t",skip=5,fill=TRUE,header=TRUE)
proteinClass = proteinClass[proteinClass[,3]>1,]
proteinClass$pvalue = as.numeric(proteinClass[,5]=="+")*proteinClass[,7] + as.numeric(proteinClass[,5]=="-")*(1-proteinClass[,7])
proteinClass$fdr = p.adjust(proteinClass$pvalue,method="BH")

colnames(paths) = colnames(proteinClass) = c("PANTHER Category","bkgrd.Num","pink.Num","expected","direction","foldEnrich","two.sided.p","pvalue","fdr")

printPaths = paths[paths$pvalue<0.05,c("PANTHER Category","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]
printProteinClass = proteinClass[proteinClass$pvalue<0.05,c("PANTHER Category","bkgrd.Num","pink.Num","expected","foldEnrich","pvalue","fdr")]

printPaths$fdr = sprintf("%.3f",round(printPaths$fdr,3))
printProteinClass$fdr = sprintf("%.3f",round(printProteinClass$fdr,3))

printPaths$pvalue = sprintf("%.4f",round(printPaths$pvalue,4))
printProteinClass$pvalue = sprintf("%.4f",round(printProteinClass$pvalue,4))
```

### Enriched PANTHER Pathways in Pink Module with Nos1 (p<0.05)
```{r, results='asis',echo=FALSE}
kable(printPaths,align=rep("c",ncol(printPaths)))
```

### Enriched PANTHER Protein Classes in Pink Module with Nos1 (p<0.05)
```{r, results='asis',echo=FALSE}
kable(printProteinClass,align=rep("c",ncol(printProteinClass)))
```









