BNLx and SHR Brain (polyA) Transcriptome Reconstruction
========================================================

polyA+-selected RNA Reconstruction
--------------------


###1.  Strain-Specific Transcriptome Reconstruction (Ensembl-Guided) - 

```
qsub -q smp -N BNLx.polyA /storage/saba/BNLx.SHR.Brain.polyARNA/programs/reconBNLx.brain.polyA.sh
qsub -q smp -N SHR.polyA /storage/saba/BNLx.SHR.Brain.polyARNA/programs/reconSHR.brain.polyA.sh

qsub -q smp -N BNLx.polyA /storage/saba/BNLx.SHR.Brain.polyARNA/programs/reconBNLx.brain.polyA.naive.sh
qsub -q smp -N SHR.polyA /storage/saba/BNLx.SHR.Brain.polyARNA/programs/reconSHR.brain.polyA.naive.sh

```

```{r,echo=FALSE}
rm(list=ls())
options(stringsAsFactors=FALSE)
setwd("~/Documents/BNLx.SHR/RNA-Seq.Brain/polyA/reconstruction")
printNow=FALSE

#######################
##  local functions  ##
#######################

# format GTF files to include transcript ids, gene ids, and fpkm values
gtfFormat = function(data,FPKM=TRUE){
  data$transcript_id = unlist(lapply(strsplit(data$V9,split=";",fixed=TRUE),function(a) gsub(" transcript_id ","",a[grep("transcript",a)]))) 
  data$gene_id = unlist(lapply(strsplit(data$V9,split=";",fixed=TRUE),function(a) gsub("gene_id ","",a[grep("gene",a)]))) 
  if(FPKM) data$fpkm = as.numeric(unlist(lapply(strsplit(data$V9,split=";",fixed=TRUE),function(a) gsub(" FPKM ","",a[grep("FPKM",a)]))))
  return(data)
}

checkSame = function(x){
  sameGene = FALSE
  if(x["start.x"]==x["start.y"] | x["stop.x"]==x["stop.y"]) sameGene = TRUE
  if(!is.na(x["exonJunct.x"]) & !is.na(x["exonJunct.y"]) & sum(duplicated(c(unlist(strsplit(x["exonJunct.x"],split=",")),unlist(strsplit(x["exonJunct.y"],split=",")))))>0) sameGene = TRUE
  return(sameGene)
}

# turn gtf file into one record per transcript

##  create one per transcript ##

gtfToOnePer = function(gtf){
  start = aggregate(gtf$V4,by=list(gtf$transcript_id),function(a) paste(a,collapse=","))
  stop = aggregate(gtf$V5,by=list(gtf$transcript_id),function(a) paste(a,collapse=","))
  transcripts = gtf[!duplicated(gtf[,c("transcript_id","V1")]),c("transcript_id","V1","V6","V7")]
  onePerTrans = merge(start,stop,by=1)
  onePerTrans = merge(transcripts,onePerTrans,by=1)

  onePerTrans$exonNum = unlist(lapply(strsplit(onePerTrans$x.x,split=",",fixed=TRUE),length))
  onePerTrans$start = as.integer(unlist(lapply(strsplit(onePerTrans$x.x,split=",",fixed=TRUE),function(a) min(as.numeric(a)))))
  onePerTrans$stop = as.integer(unlist(lapply(strsplit(onePerTrans$x.y,split=",",fixed=TRUE),function(a) max(as.numeric(a)))))
  return(onePerTrans)
}



```

```{r,echo=FALSE}
options(stringsAsFactors=FALSE)

shr = read.table(file = "SHR.Brain.PolyA.ensemblGuided.rn6/transcripts.gtf",sep="\t",header=FALSE)
shr = gtfFormat(shr)
shr.v2 = gtfToOnePer(shr[shr$V3=="exon",])
shr.dups = merge(shr.v2[duplicated(shr.v2[,c("V1","V7","x.x","x.y","exonNum","start","stop")]),c("V1","V7","x.x","x.y","exonNum","start","stop")],shr.v2,by=c("V1","V7","x.x","x.y","exonNum","start","stop"))
shr.dups.remove = shr.dups$transcript_id[grep("CUFF",shr.dups$transcript_id)]

shr = shr[!(shr$transcript_id %in% shr.dups.remove),]


bnlx = read.table(file = "BNLx.Brain.PolyA.ensemblGuided.rn6/transcripts.gtf",sep="\t",header=FALSE)
bnlx = gtfFormat(bnlx)
bnlx.v2 = gtfToOnePer(bnlx[bnlx$V3=="exon",])
bnlx.dups = merge(bnlx.v2[duplicated(bnlx.v2[,c("V1","V7","x.x","x.y","exonNum","start","stop")]),c("V1","V7","x.x","x.y","exonNum","start","stop")],bnlx.v2,by=c("V1","V7","x.x","x.y","exonNum","start","stop"))
bnlx.dups.remove = bnlx.dups$transcript_id[grep("CUFF",bnlx.dups$transcript_id)]

bnlx = bnlx[!(bnlx$transcript_id %in% bnlx.dups.remove),]


shr.transcripts = shr[shr$V3=="transcript",]
shr.transcripts = shr.transcripts[shr.transcripts$fpkm>0,]

bnlx.transcripts = bnlx[bnlx$V3=="transcript",]
bnlx.transcripts = bnlx.transcripts[bnlx.transcripts$fpkm>0,]

shr.transcripts$transcript_id = gsub("CUFF","SHR",shr.transcripts$transcript_id)
bnlx.transcripts$transcript_id = gsub("CUFF","BNLx",bnlx.transcripts$transcript_id)
```

Complete SHR Transcriptome:  
`r prettyNum(length(unique(shr.transcripts$gene_id)),big.mark=",")` genes  
`r prettyNum(nrow(shr.transcripts),big.mark=",")` transcripts  
`r prettyNum(sum(grepl("ENS",shr.transcripts$transcript_id)),big.mark=",")` Ensembl transcripts  
`r prettyNum(sum(!grepl("ENS",shr.transcripts$transcript_id)),big.mark=",")` novel transcripts  

Complete BNLx Transcriptome:  
`r prettyNum(length(unique(bnlx.transcripts$gene_id)),big.mark=",")` genes  
`r prettyNum(nrow(bnlx.transcripts),big.mark=",")` transcripts  
`r prettyNum(sum(grepl("ENS",bnlx.transcripts$transcript_id)),big.mark=",")` Ensembl transcripts  
`r prettyNum(sum(!grepl("ENS",bnlx.transcripts$transcript_id)),big.mark=",")` novel transcripts  

1. Limit to transcripts with greater than 50 coverage and a transcript length greater than 300 nt (not including introns)
--------------

```{r,echo=FALSE}
options(stringsAsFactors=FALSE)

#########################################################
###  Look For Overlap of High Confidence Transcripts  ###
#########################################################

shr.tracking = read.table(file = "SHR.Brain.PolyA.ensemblGuided.rn6/isoforms.fpkm_tracking",sep="\t",header=TRUE)
shr.tracking = shr.tracking[!(shr.tracking$tracking_id %in% shr.dups.remove),]

bnlx.tracking = read.table(file = "BNLx.Brain.PolyA.ensemblGuided.rn6/isoforms.fpkm_tracking",sep="\t",header=TRUE)
bnlx.tracking = bnlx.tracking[!(bnlx.tracking$tracking_id %in% bnlx.dups.remove),]

high.bnlx = bnlx.tracking[bnlx.tracking$FPKM>0,]
high.bnlx = high.bnlx[high.bnlx$length>300,]
high.bnlx = high.bnlx[high.bnlx$FPKM>0.01 & high.bnlx$coverage>0.1,]

high.shr = shr.tracking[shr.tracking$FPKM>0,]
high.shr = high.shr[high.shr$length>300,]
high.shr = high.shr[high.shr$FPKM>0.01 & high.shr$coverage>0.1,]

high.shr = high.shr[!grepl("ERCC",high.shr$locus),]
high.bnlx = high.bnlx[!grepl("ERCC",high.bnlx$locus),]

high.bnlx$transcript_id = gsub("CUFF","BNLx",high.bnlx$tracking_id)
high.shr$transcript_id = gsub("CUFF","SHR",high.shr$tracking_id)

###  Characterize  ###
gene.shr = aggregate(shr.tracking$tracking_id,by=list(gene_id=shr.tracking$gene_id),function(a) sum(grepl("ENS",a)))
high.shr = merge(high.shr,gene.shr,by="gene_id")

gene.bnlx = aggregate(bnlx.tracking$tracking_id,by=list(gene_id=bnlx.tracking$gene_id),function(a) sum(grepl("ENS",a)))
high.bnlx = merge(high.bnlx,gene.bnlx,by="gene_id")

###  Examine High Expressors  ###
high.shr = high.shr[order(high.shr$FPKM,decreasing=TRUE),]
high.bnlx = high.bnlx[order(high.bnlx$FPKM,decreasing=TRUE),]
```

High Confidence SHR Transcripts:  
`r prettyNum(length(unique(high.shr$gene_id)),big.mark=",")` genes  
`r prettyNum(nrow(high.shr),big.mark=",")` transcripts  
`r prettyNum(sum(grepl("ENS",high.shr$transcript_id)),big.mark=",")` Ensembl transcripts  
`r prettyNum(sum(!grepl("ENS",high.shr$transcript_id) & high.shr$x>0),big.mark=",")` novel transcripts of Ensembl genes  
`r prettyNum(sum(!grepl("ENS",high.shr$transcript_id) & high.shr$x==0),big.mark=",")` novel transcripts/genes   


High Confidence BNLx Transcripts:  
`r prettyNum(length(unique(high.bnlx$gene_id)),big.mark=",")` genes  
`r prettyNum(nrow(high.bnlx),big.mark=",")` transcripts  
`r prettyNum(sum(grepl("ENS",high.bnlx$transcript_id)),big.mark=",")` Ensembl transcripts  
`r prettyNum(sum(!grepl("ENS",high.bnlx$transcript_id) & high.bnlx$x>0),big.mark=",")` novel transcripts of Ensembl genes  
`r prettyNum(sum(!grepl("ENS",high.bnlx$transcript_id) & high.bnlx$x==0),big.mark=",")` novel transcripts/genes   



2. Identify overlap between novel transcripts in strain-specific reconstructions
---------------------------------
```{r,echo=FALSE}
high.combined = merge(high.bnlx,high.shr,by=c("transcript_id","locus","length"),all=TRUE)
high.combined = high.combined[order(high.combined$locus),c("transcript_id","locus","length","coverage.x","FPKM.x","coverage.y","FPKM.y")]

bnlx.bed = bnlx.transcripts[bnlx.transcripts$transcript_id %in% high.combined$transcript_id,]
bnlx.bed$V6 = 1000

shr.bed = shr.transcripts[shr.transcripts$transcript_id %in% high.combined$transcript_id,]
shr.bed$V6 = 1000

bnlx.bed = bnlx.bed[,c("V1","V4","V5","V6","V7","transcript_id")]
shr.bed = shr.bed[,c("V1","V4","V5","V6","V7","transcript_id")]

if(printNow) write.table(shr.bed,file="tmp/shr.bed",sep="\t",row.names=FALSE,col.names=FALSE,quote=FALSE)
if(printNow) write.table(bnlx.bed,file="tmp/bnlx.bed",sep="\t",row.names=FALSE,col.names=FALSE,quote=FALSE)

```

BEDtools version = bedtools v2.24.0
```
cd ~/Documents/BNLx.SHR/RNA-Seq.Brain/polyA/reconstruction/tmp
intersectBed -a shr.bed -b bnlx.bed -wo > overlap.brain.polyA.byStrain.txt
```

3.  Compare novel transcripts across strain-specific transcriptomes BN-Lx and SHR 
------------------------------------------------------
Two transcripts were "merged" into one transcripts if:  
* they are both assigned to the same strand or one/both did not have a strand designation
* if 1) all exon starts and exon stops matched OR 2) all exon junctions matched, transcription start and stop sites could differ OR 3) two one-exon transcripts with transcription start sites within 100 bp of each other and transcription stop sites within 100 bp of each other

```{r,eval=TRUE,echo=FALSE}
options(stringsAsFactors=FALSE)
setwd("~/Documents/BNLx.SHR/RNA-Seq.Brain/polyA/reconstruction")
printNow=FALSE

shr$transcript_id = gsub("CUFF","SHR",shr$transcript_id)
shr.gtf <- shr[(shr$transcript_id %in% high.combined$transcript_id) & shr$V3=="exon",]
shr.gtf <- shr.gtf[order(shr.gtf$transcript_id,shr.gtf$V4,shr.gtf$V5),]

bnlx$transcript_id = gsub("CUFF","BNLx",bnlx$transcript_id)
bnlx.gtf <- bnlx[(bnlx$transcript_id %in% high.combined$transcript_id) & bnlx$V3=="exon",]
bnlx.gtf <- bnlx.gtf[order(bnlx.gtf$transcript_id,bnlx.gtf$V4,bnlx.gtf$V5),]

gtf = rbind(shr.gtf,bnlx.gtf)
gtf = gtf[,!(colnames(gtf) %in% c("gene_id","fpkm","V9"))]
gtf = gtf[!duplicated(gtf),]
gtf <- gtf[order(gtf$transcript_id,gtf$V4,gtf$V5),]

##  create one per transcript ##
onePerTrans = gtfToOnePer(gtf)

#####################
overlap <- read.table(file="tmp/overlap.brain.polyA.byStrain.txt",sep="\t",header=FALSE)
colnames(overlap) <- c(paste(c("chr","start","stop","V5","strand","transcript_id"),rep(c("SHR","BNLx"),each=6),sep="."),"overlap")

## remove records that are overlap of the same Ensembl transcript
overlap = overlap[overlap$transcript_id.SHR!=overlap$transcript_id.BNLx,]

## add exon information to overlaps
setSHR = setBNLx = onePerTrans[,c("transcript_id","x.x","x.y","exonNum")]
colnames(setSHR) = c("transcript_id.SHR","exonStarts.SHR","exonStops.SHR","exonNum.SHR")
colnames(setBNLx) = c("transcript_id.BNLx","exonStarts.BNLx","exonStops.BNLx","exonNum.BNLx")

overlap = merge(overlap,setSHR,by="transcript_id.SHR")
overlap = merge(overlap,setBNLx,by="transcript_id.BNLx")

overlap$ID = c(1:nrow(overlap))

##  identify exon junction

findJunct <- function(starts,stops){
  start <- strsplit(starts,split=",",fixed=TRUE)[[1]]
  stop <- strsplit(stops,split=",",fixed=TRUE)[[1]]
  junct <- NA
  if(length(start)>1) junct <- paste(paste(stop[-length(stop)],start[-1],sep="//"),collapse=",")
  return(junct)
  }
  
overlap$exonJunct.SHR = apply(overlap[,c("exonStarts.SHR","exonStops.SHR")],1,function(a) findJunct(a[1],a[2]))
overlap$exonJunct.BNLx = apply(overlap[,c("exonStarts.BNLx","exonStops.BNLx")],1,function(a) findJunct(a[1],a[2]))

## remove overlap that is on the opposite strand
overlap <- overlap[overlap$strand.SHR=="." | overlap$strand.BNLx=="." | overlap$strand.SHR==overlap$strand.BNLx,]

## remove overlap where coding regions don't overlap  ##
intronCheck =c()
for(i in 1:nrow(overlap)){
  input = overlap[i,]
  out = sum(sum(unlist(lapply(strsplit(strsplit(input[1,"exonJunct.BNLx"],split=",")[[1]],split="//"),function(a) input[1,"start.SHR"]>a[1] & input[1,"stop.SHR"]<a[2]))),sum(unlist(lapply(strsplit(strsplit(input[1,"exonJunct.SHR"],split=",")[[1]],split="//"),function(a) input[1,"start.BNLx"]>a[1] & input[1,"stop.BNLx"]<a[2]))),na.rm=TRUE)
  intronCheck = c(intronCheck,out)
  }
  
overlap = overlap[intronCheck==0,]

## transcripts without any overlap - 28,354 transcripts
noOverlap = onePerTrans[!(onePerTrans$transcript_id %in% unique(c(overlap$transcript_id.SHR,overlap$transcript_id.BNLx))),]

##  perfect overlap - 119 transcripts
perfect = overlap[overlap$exonStarts.SHR==overlap$exonStarts.BNLx & overlap$exonStops.SHR==overlap$exonStops.BNLx,]

##  identify exon junction matches - 4,714 transcripts
reduced = overlap[!(overlap$ID %in% perfect$ID),]
exonJunctMatch <- reduced[!is.na(reduced$exonJunct.SHR) & !is.na(reduced$exonJunct.BNLx) & reduced$exonJunct.SHR==reduced$exonJunct.BNLx,]

##  examine one-exon transcripts - 12,404 overlaps between one-exon transcripts
reduced = overlap[!(overlap$ID %in% c(perfect$ID,exonJunctMatch$ID)),]
oneExons = reduced[reduced$exonNum.SHR==1 & reduced$exonNum.BNLx==1,]

## close match one-exon transcripts - 1,616 transcripts
closeMatch.oneExon = oneExons[abs(oneExons$start.SHR - oneExons$start.BNLx)<100 & abs(oneExons$stop.SHR - oneExons$stop.BNLx)<100, ]

##  Check for overlap between renamed  ##
matched = as.data.frame(rbind(perfect,exonJunctMatch,closeMatch.oneExon))

bDups = matched[matched$transcript_id.BNLx %in% matched$transcript_id.BNLx[duplicated(matched$transcript_id.BNLx)],]
bDups$mismatch = abs(bDups$start.SHR - bDups$start.BNLx) + abs(bDups$stop.SHR - bDups$stop.BNLx)
bDups = bDups[order(bDups$transcript_id.BNLx,bDups$mismatch),]
bDups.remove = bDups[duplicated(bDups$transcript_id.BNLx),"ID"]

matched = matched[!(matched$ID %in% bDups.remove),]

sDups = matched[matched$transcript_id.SHR %in% matched$transcript_id.SHR[duplicated(matched$transcript_id.SHR)],]
sDups$mismatch = abs(sDups$start.SHR - sDups$start.BNLx) + abs(sDups$stop.SHR - sDups$stop.BNLx)
sDups = sDups[order(sDups$transcript_id.SHR,sDups$mismatch),]
sDups.remove = sDups[duplicated(sDups$transcript_id.SHR),"ID"]

matched = matched[!(matched$ID %in% sDups.remove),]

matched.dups = matched[(matched$transcript_id.BNLx %in% matched$transcript_id.SHR) | (matched$transcript_id.SHR %in% matched$transcript_id.BNLx),]

matched.dups.v1 = matched.dups[grep("ENS",matched.dups$transcript_id.BNLx),]
colnames(matched.dups.v1) = gsub("BNLx","Ens",colnames(matched.dups.v1)) 
  
matched.dups.v2 = matched.dups[grep("ENS",matched.dups$transcript_id.SHR),]
colnames(matched.dups.v2) = gsub("SHR","Ens",colnames(matched.dups.v2))

matched.dups.toAdd = merge(matched.dups.v1,matched.dups.v2,by=colnames(matched.dups.v1)[grep("Ens",colnames(matched.dups.v1))])

matched.dups.toAdd$transcript_id = paste("ensMatched",c(1:nrow(matched.dups.toAdd)),sep=".")
matched.dups.toAdd$gene_id = NA
matched.dups.toAdd$V1 = matched.dups.toAdd$chr.SHR
matched.dups.toAdd$V6 = 1000
matched.dups.toAdd$V7 = matched.dups.toAdd$strand.SHR
matched.dups.toAdd$start = apply(matched.dups.toAdd[,c("start.BNLx","start.SHR","start.Ens")],1,min)
matched.dups.toAdd$stop = apply(matched.dups.toAdd[,c("stop.BNLx","stop.SHR","start.Ens")],1,max)
matched.dups.toAdd$exonNum = matched.dups.toAdd$exonNum.SHR

matched.dups.toAdd$x.x = NA
matched.dups.toAdd$x.x[matched.dups.toAdd$start==matched.dups.toAdd$start.SHR] = matched.dups.toAdd$exonStarts.SHR[matched.dups.toAdd$start==matched.dups.toAdd$start.SHR]
matched.dups.toAdd$x.x[matched.dups.toAdd$start==matched.dups.toAdd$start.BNLx] = matched.dups.toAdd$exonStarts.BNLx[matched.dups.toAdd$start==matched.dups.toAdd$start.BNLx]
matched.dups.toAdd$x.x[matched.dups.toAdd$start==matched.dups.toAdd$start.Ens] = matched.dups.toAdd$exonStarts.Ens[matched.dups.toAdd$start==matched.dups.toAdd$start.Ens]

matched.dups.toAdd$x.y = NA
matched.dups.toAdd$x.y[matched.dups.toAdd$stop==matched.dups.toAdd$stop.SHR] = matched.dups.toAdd$exonStops.SHR[matched.dups.toAdd$stop==matched.dups.toAdd$stop.SHR]
matched.dups.toAdd$x.y[matched.dups.toAdd$stop==matched.dups.toAdd$stop.BNLx] = matched.dups.toAdd$exonStops.BNLx[matched.dups.toAdd$stop==matched.dups.toAdd$stop.BNLx]
matched.dups.toAdd$x.y[matched.dups.toAdd$stop==matched.dups.toAdd$stop.Ens] = matched.dups.toAdd$exonStops.Ens[matched.dups.toAdd$stop==matched.dups.toAdd$stop.Ens]

matched.undup = matched[!(matched$transcript_id.BNLx %in% matched.dups$transcript_id.BNLx) & !(matched$transcript_id.SHR %in% matched.dups$transcript_id.SHR),]

##  create new transcript for matched transcripts - 2,245 matches
matched.undup$transcript_id = paste("matched",c(1:nrow(matched.undup)),sep=".")
matched.undup$gene_id = NA
matched.undup$V1 = matched.undup$chr.SHR
matched.undup$V6 = 1000
matched.undup$V7 = matched.undup$strand.SHR
matched.undup$start = apply(matched.undup[,c("start.BNLx","start.SHR")],1,min)
matched.undup$stop = apply(matched.undup[,c("stop.BNLx","stop.SHR")],1,max)
matched.undup$exonNum = matched.undup$exonNum.SHR

matched.undup$x.x = NA
matched.undup$x.x[matched.undup$start==matched.undup$start.SHR] = matched.undup$exonStarts.SHR[matched.undup$start==matched.undup$start.SHR]
matched.undup$x.x[matched.undup$start==matched.undup$start.BNLx] = matched.undup$exonStarts.BNLx[matched.undup$start==matched.undup$start.BNLx]

matched.undup$x.y = NA
matched.undup$x.y[matched.undup$stop==matched.undup$stop.SHR] = matched.undup$exonStops.SHR[matched.undup$stop==matched.undup$stop.SHR]
matched.undup$x.y[matched.undup$stop==matched.undup$stop.BNLx] = matched.undup$exonStops.BNLx[matched.undup$stop==matched.undup$stop.BNLx]

matched.toAdd = rbind(matched.undup[,colnames(onePerTrans)],matched.dups.toAdd[,colnames(onePerTrans)])

new.transcripts = onePerTrans[!(onePerTrans$transcript_id %in% c(matched$transcript_id.SHR,matched$transcript_id.BNLx)),]
new.transcripts = rbind(new.transcripts,matched.toAdd)

if(printNow) write.table(new.transcripts[,c("V1","start","stop","transcript_id","V6","V7")],file="RNA-Seq.Brain.polyA/reconstruction/tmp/combined.brain.polyA.10Oct14.bed",sep="\t",quote=FALSE,row.names=FALSE,col.names=FALSE)
```

Combined Transcriptome:  
* `r prettyNum(nrow(new.transcripts),big.mark=",")` transcripts
* `r prettyNum(sum(grepl("matched",new.transcripts$transcript_id)),big.mark=",")` novel transcripts identified in both strains
* `r prettyNum(sum(grepl("BNLx",new.transcripts$transcript_id)),big.mark=",")` novel transcripts only identified in BNLx
* `r prettyNum(sum(grepl("SHR",new.transcripts$transcript_id)),big.mark=",")` novel transcripts only identified in SHR
* `r prettyNum(sum(grepl("ENS",new.transcripts$transcript_id)),big.mark=",")` Ensembl transcripts
* `r prettyNum(sum(grepl("ens",new.transcripts$transcript_id)),big.mark=",")` novel transcripts very similar to Ensembl


4.  Identify overlap between transcripts in order to identify transcripts from the same gene
------------------------------------------------------

BEDtools Version = bedtools v2.19.1
```
export PATH=/usr/local/bedtools2/bin:$PATH
cd /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/reconstruction/tmp
intersectBed -a combined.brain.polyA.10Oct14.bed -b combined.brain.polyA.10Oct14.bed -wo > overlap.brain.polyA.forGeneID.txt
```

5.  Match transcripts to genes and create new GTF file  
------------------------------------------------------

Two transcripts are from the same gene if:  
* Their transcription start sites matched exactly OR
* Their transcription stop sites matched exactly OR
* At least one exon-exon junction matched exactly

```{r,eval=FALSE,echo=FALSE}
findJunct <- function(starts,stops){
  start <- strsplit(starts,split=",",fixed=TRUE)[[1]]
  stop <- strsplit(stops,split=",",fixed=TRUE)[[1]]
  junct <- NA
  if(length(start)>1) junct <- paste(paste(stop[-length(stop)],start[-1],sep="//"),collapse=",")
  return(junct)
  }

new.transcripts$exonJunct = apply(new.transcripts[,c("x.x","x.y")],1,function(a) findJunct(a[1],a[2]))

overlap = read.table(file="RNA-Seq.Brain.polyA/reconstruction/tmp/overlap.brain.polyA.forGeneID.txt",sep="\t",header=FALSE)
overlap = overlap[overlap$V4!=overlap$V10,]
overlap = overlap[overlap$V6=="." | overlap$V12=="." | overlap$V6==overlap$V12,]

overlap$pair = apply(overlap,1,function(a) paste(sort(a[c("V4","V10")]),collapse="//"))
overlap = overlap[!duplicated(overlap$pair),]

overlap2 = merge(overlap[,c("V4","V10")],new.transcripts[,c("transcript_id","start","stop","exonJunct")],by.x="V4",by.y="transcript_id")
overlap2 = merge(overlap2,new.transcripts[,c("transcript_id","start","stop","exonJunct")],by.x="V10",by.y="transcript_id")

match = apply(overlap2,1,checkSame)
pairs = overlap2[match,c("V4","V10")]

##  identify genes with multiple transcripts and create gene identifier  ##
library("igraph")
library("ggm")

edges <- pairs
graph.polyA <- graph.data.frame(edges)
graph.polyA <- igraph.to.graphNEL(graph.polyA)

compList <- connectedComp(graph.polyA)

geneNames = unlist(lapply(compList,function(a) paste(a[grep("ENS",a)],collapse="//")))
geneNames[geneNames==""] = paste("GENE.",c(1:sum(geneNames=="")),sep="")
geneNames = data.frame(geneNames)
geneNames$order = c(1:nrow(geneNames))

names(compList) <- paste("GENE",c(1:length(compList)),sep=".")

multiTrans <- sapply(names(compList),function(a) cbind(gene_id.new=a,transcript_id.new=paste(a,c(1:length(compList[[a]])),sep="."),id=compList[[a]]))
multiTrans <- data.frame(do.call(rbind,multiTrans))

singleTrans <- data.frame(id=new.transcripts$transcript_id[!(new.transcripts$transcript_id %in% multiTrans$id)])
singleTrans$gene_id.new = paste("GENE",c((length(compList)+1):(length(compList)+nrow(singleTrans))),sep=".")
singleTrans$transcript_id.new = paste(singleTrans$gene_id.new,1,sep=".")

ids = rbind(multiTrans,singleTrans[,colnames(multiTrans)])

convertID = function(x,y){
if(!grepl("ENS",x)) y = gsub("GENE",unlist(lapply(strsplit(x,split=".",fixed=TRUE),function(a) a[1])),y)
if(grepl("ENS",x)) y = x
return(y)
}

ids$alt.id = apply(ids,1,function(a) convertID(a["id"],a["transcript_id.new"]))

updated <- merge(new.transcripts,ids,by.x="transcript_id",by.y="id")

##  Create New GTF File  ##

byExon = data.frame(do.call(rbind,apply(updated,1,function(tmp) data.frame(alt.id = as.character(tmp["alt.id"]),V4=unlist(strsplit(as.character(tmp[5]),split=",")),V5=unlist(strsplit(as.character(tmp[6]),split=","))))))

gtf = merge(updated,byExon,by="alt.id")
gtf$V2 = "CuffLinks"
gtf$V3 = "exon"
gtf$V8 = "."
gtf$V9 = paste('gene_id "',gtf$gene_id.new,'"; transcript_id "',gtf$alt.id,'"; original "',gtf$transcript_id,'";',sep="")

gtf = gtf[,paste("V",c(1:9),sep="")]
gtf = gtf[!duplicated(gtf),]

if(printNow) write.table(gtf,file="RNA-Seq.Brain.polyA/reconstruction/reconstruct.brain.polyA.10Oct14.v1.gtf",row.names=FALSE,col.names=FALSE,quote=FALSE,sep="\t")
```

6. Quantitate Combined Transcriptome (1st Iteration)
---------------------------------
```
#on PhenoGen server
cufflinks -u --seed 5645 -p 25 --library-type fr-unstranded -b /data2/saba/index/BNLx_rn5_wSpikesAndM.fa -G /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/reconstruction/reconstruct.brain.polyA.10Oct14.v1.gtf -o /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/quantitation/recon.Brain.polyA.v1.BNLx /hi-seq/BNLx.SHR.Brain-polyARNA/alignedReads/BNLx.brain.polyA.bam
cufflinks -u --seed 19885 -p 25 --library-type fr-unstranded -b /data2/saba/index/SHR_rn5_wSpikesAndM.fa -G /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/reconstruction/reconstruct.brain.polyA.10Oct14.v1.gtf -o /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/quantitation/recon.Brain.polyA.v1.SHR /hi-seq/BNLx.SHR.Brain-polyARNA/alignedReads/SHR.brain.polyA.bam
```

```{r,echo=FALSE}
#rm(list=ls())
options(stringsAsFactors=FALSE)
setwd("/Volumes/saba/BNLx.SHR/RNA-Seq.Brain.polyA/")
printNow=FALSE

shr = read.table(file="quantitation/recon.Brain.polyA.v1.SHR/isoforms.fpkm_tracking",sep="\t",header=TRUE)
bnlx = read.table(file="quantitation/recon.Brain.polyA.v1.BNLx/isoforms.fpkm_tracking",sep="\t",header=TRUE)

## started wtih 19,793 transcripts  ##

quant = merge(shr[,c("tracking_id","gene_id","locus","length","coverage","FPKM")],bnlx[,c("tracking_id","gene_id","locus","length","coverage","FPKM")],by=c("tracking_id","gene_id","locus","length"),suffixes=c(".shr",".bnlx"))

##  Keep Only Transcripts with Greater Than 50 Read Coverage ##
keepers = quant[quant$coverage.shr>50 | quant$coverage.bnlx>50,]
#nrow(keepers) / nrow(quant) 

## 17,284 (87%) transcripts with greater than 50 read coverage

gtf.old = gtfFormat(read.table(file="quantitation/recon.Brain.polyA.v1.SHR/transcripts.gtf",header=FALSE,sep="\t"))

## rename transcripts based on FPKM  ##
keepers = keepers[order(keepers$gene_id,rowMeans(keepers[,grep("FPKM",colnames(keepers))]),decreasing=TRUE),]
newID = data.frame(do.call(rbind,aggregate(keepers$tracking_id,by=list(gene_id=keepers$gene_id),function(a) cbind(a,paste(unlist(lapply(strsplit(a,split=".",fixed=TRUE),function(b) paste(b[-3],collapse="."))),1:length(a),sep=".")))$x))
colnames(newID) = c("transcript_id","transcript_id.new")

## remove suffix on Ensembl transcripts
newID$transcript_id.new[grep("ENSRN",newID$transcript_id.new)] = gsub("\\..*","",newID$transcript_id.new[grep("ENSRN",newID$transcript_id.new)])
  
##  reduce GTF file  ##
new.gtf = merge(gtf.old,newID,by="transcript_id")
new.gtf$V9 = paste('gene_id "',new.gtf$gene_id,'"; transcript_id "',new.gtf$transcript_id.new,'";',sep="")
new.gtf = new.gtf[,paste("V",c(1:9),sep="")]

if(printNow) write.table(new.gtf,file="reconstruction/reconstruct.brain.polyA.13Oct14.v2.gtf",row.names=FALSE,col.names=FALSE,quote=FALSE,sep="\t")
````

In the first iteration, `r prettyNum(nrow(keepers),big.mark=",")` out of `r prettyNum(nrow(bnlx),big.mark=",")` transcripts were retained (`r prettyNum(length(unique(keepers$gene_id)),big.mark=",")` out of `r prettyNum(length(unique(bnlx$gene_id)),big.mark=",")` genes).


7. Quantitate Combined Transcriptome (2nd Iteration)
---------------------------------
```
cufflinks -u --seed 65498 -p 25 --library-type fr-unstranded -b /data2/saba/index/BNLx_rn5_wSpikesAndM.fa -G /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/reconstruction/reconstruct.brain.polyA.13Oct14.v2.gtf -o /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/quantitation/recon.Brain.polyA.v2.BNLx /hi-seq/BNLx.SHR.Brain-polyARNA/alignedReads/BNLx.brain.polyA.bam
cufflinks -u --seed 56946 -p 25 --library-type fr-unstranded -b /data2/saba/index/SHR_rn5_wSpikesAndM.fa -G /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/reconstruction/reconstruct.brain.polyA.13Oct14.v2.gtf -o /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/quantitation/recon.Brain.polyA.v2.SHR /hi-seq/BNLx.SHR.Brain-polyARNA/alignedReads/SHR.brain.polyA.bam
```

```{r,echo=FALSE,eval=TRUE}
#rm(list=ls())
options(stringsAsFactors=FALSE)
setwd("/Volumes/saba/BNLx.SHR/RNA-Seq.Brain.polyA/")
printNow=FALSE

shr = read.table(file="quantitation/recon.Brain.polyA.v2.SHR/isoforms.fpkm_tracking",sep="\t",header=TRUE)
bnlx = read.table(file="quantitation/recon.Brain.polyA.v2.BNLx/isoforms.fpkm_tracking",sep="\t",header=TRUE)

## started wtih 17,284 transcripts  ##

quant = merge(shr[,c("tracking_id","gene_id","locus","length","coverage","FPKM")],bnlx[,c("tracking_id","gene_id","locus","length","coverage","FPKM")],by=c("tracking_id","gene_id","locus","length"),suffixes=c(".shr",".bnlx"))

##  Keep Only Transcripts with Greater Than 50 Read Coverage ##
keepers = quant[quant$coverage.shr>50 | quant$coverage.bnlx>50,]
#nrow(keepers) / nrow(quant) 

## 17,264 (99.9%) transcripts with greater than 50 read coverage

gtf.old = read.table(file="quantitation/recon.Brain.polyA.v2.SHR/transcripts.gtf",header=FALSE,sep="\t")
gtf.old = gtfFormat(gtf.old)

## rename transcripts based on TPM  ##
keepers = keepers[order(keepers$gene_id,rowMeans(keepers[,grep("FPKM",colnames(keepers))]),decreasing=TRUE),]
newID = data.frame(do.call(rbind,aggregate(keepers$tracking_id,by=list(gene_id=keepers$gene_id),function(a) cbind(a,paste(unlist(lapply(strsplit(a,split=".",fixed=TRUE),function(b) paste(b[-3],collapse="."))),1:length(a),sep=".")))$x))
colnames(newID) = c("transcript_id","transcript_id.new")

## remove suffix on Ensembl transcripts
newID$transcript_id.new[grep("ENSRN",newID$transcript_id.new)] = gsub("\\..*","",newID$transcript_id.new[grep("ENSRN",newID$transcript_id.new)])
  
##  reduce GTF file  ##
new.gtf = merge(gtf.old,newID,by="transcript_id")
new.gtf$V9 = paste('gene_id "',new.gtf$gene_id,'"; transcript_id "',new.gtf$transcript_id.new,'";',sep="")
new.gtf = new.gtf[,paste("V",c(1:9),sep="")]

if(printNow) write.table(new.gtf,file="reconstruction/reconstruct.brain.polyA.14Oct14.FINAL.gtf",row.names=FALSE,col.names=FALSE,quote=FALSE,sep="\t")
````

In the second (AND FINAL) iteration, `r prettyNum(nrow(keepers),big.mark=",")` out of `r prettyNum(nrow(bnlx),big.mark=",")` transcripts were retained (`r prettyNum(length(unique(keepers$gene_id)),big.mark=",")` out of `r prettyNum(length(unique(bnlx$gene_id)),big.mark=",")` genes).

8. Quantitate Final Transcriptome For Annotation
---------------------------------
```
cufflinks -u --seed 5645 -p 25 --library-type fr-unstranded -b /data2/saba/index/BNLx_rn5_wSpikesAndM.fa -G /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/reconstruction/reconstruct.brain.polyA.14Oct14.FINAL.gtf -o /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/quantitation/recon.Brain.polyA.FINAL.BNLx /hi-seq/BNLx.SHR.Brain-polyARNA/alignedReads/BNLx.brain.polyA.bam
cufflinks -u --seed 98135 -p 25 --library-type fr-unstranded -b /data2/saba/index/SHR_rn5_wSpikesAndM.fa -G /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/reconstruction/reconstruct.brain.polyA.14Oct14.FINAL.gtf -o /data2/saba/BNLx.SHR/RNA-Seq.Brain.polyA/quantitation/recon.Brain.polyA.FINAL.SHR /hi-seq/BNLx.SHR.Brain-polyARNA/alignedReads/SHR.brain.polyA.bam

```

