RI Panel - Batch 2 (Brain) Total RNA RNA-Seq
========================================================

* 1 sample per strain (SHR, BXH2, HXB10, HXB15, HXB18, HXB1, HXB20, HXB21, HXB22, HXB23, HXB24, HXB29, HXB31, HXB3, HXB4, HXB5, SHR)
* tissues received from Dr. Michal Pravenec at Academy of Sciences of the Czech Republic, Prague, Czech Republic
* ribosomal RNA depleted total RNA
* 2X100 reads using the stranded protocol
* sequencing done by Beckman Coulter Genomics - received September 24, 2014
  * labeled UColorado Tabakoff Batch 5 by Beckman
* synthetic spike-in were added to each sample 
  * HXB1_1, HXB3_1, HXB4_1, HXB15_1, HXB23_1, HXB24_1, HXB31_1, and SHR_1 have mix 1
  * HXB5_1, HXB10_1, HXB18_1, HXB20_1, HXB21_1, HXB22_1, HXB29_1, and BXH2_1 have mix 2
  * 4 mL of 1:100 diluted spikes were added after RNA extraction but before rRNA-depletion
  * Information regarding the spike-ins can be found here: http://products.invitrogen.com/ivgn/product/4456739

1. Unzip Files - DONE
---------------
```
qsub -q smp /storage/saba/Rat.RI.Brain.totalRNA.batch2/programs/unzip.yucca.sh
```

2. Determine number of reads sent for each sample - DONE
-------------------------------------------------
```
qsub -q smp /storage/saba/Rat.RI.Brain.totalRNA.batch2/programs/countRawReads.brain.batch2.sh
```

```{r,echo=FALSE}
rm(list=ls())
options(stringsAsFactors=FALSE)
setwd("~/Documents/NextGenSeq/Rat.RI.Brain.totalRNA.batch2/")
rawCounts = read.table(file="data/rawReadCounts.25Sep14.txt",sep=" ",header=FALSE,fill=TRUE)

rawCounts$readFrag = as.numeric(rawCounts$V3)
rawCounts$file = unlist(lapply(strsplit(rawCounts$V1,split="/",fixed=TRUE),function(a) a[length(a)]))
rawCounts$strain = unlist(lapply(strsplit(rawCounts$file,split="_",fixed=TRUE),function(a) a[1]))
rawCounts$sampleNum = unlist(lapply(strsplit(rawCounts$file,split="_",fixed=TRUE),function(a) gsub("brain","",a[2])))
rawCounts$sample = paste(rawCounts$strain,rawCounts$sampleNum,sep="_")

readFragments = aggregate(rawCounts$readFrag,by=list(sample=rawCounts$sample),sum)
readFragments$numPairedReads = prettyNum(readFragments$x/2,big.mark=",",scientific=FALSE)
readFragments$numReadFragments = prettyNum(readFragments$x,big.mark=",",scientific=FALSE)

readFragments=readFragments[,colnames(readFragments)!="x"]
forPrint = readFragments[,c("sample","numPairedReads","numReadFragments")]
colnames(forPrint) = c("sample","Number of Paired-End Reads","Number of Read Fragments")
```

Raw Reads/Read Fragments
---------------------------

```{r, results='asis',echo=FALSE}
kable(forPrint,align=rep("c",ncol(readFragments)))
```

Total Number of Paired End Reads: `r prettyNum(sum(rawCounts$readFrag)/2,big.mark=",",scientific=FALSE)`  
Total Number of Read Fragments:  `r prettyNum(sum(rawCounts$readFrag),big.mark=",",scientific=FALSE)`  
Average Number of Paired End Reads Per Sample: `r prettyNum(sum(rawCounts$readFrag/2)/length(unique(rawCounts$sample)),big.mark=",",scientific=FALSE)`  

3. Trim reads for adaptors and for quality - DONE
---------------

```
qsub -q smp /storage/saba/Rat.RI.Brain.totalRNA.batch2/programs/trimReads.brain.batch2.sh
```

4. Characterize Trimmed Reads - RUNNING
---------------
```
qsub -q smp /storage/saba/Rat.RI.Brain.totalRNA.batch2/programs/countTrimmedReads.sh
qsub -q smp /storage/saba/Rat.RI.Brain.totalRNA.batch2/programs/compressRawReads.batch2.sh

```

```{r,echo=FALSE}
options(stringsAsFactors=FALSE)
setwd("~/Documents/NextGenSeq/Rat.RI.Brain.totalRNA.batch1/")

trimmed = read.table(file="data/trimmedReadCounts.18Sep14.txt",sep="",header=FALSE)
trimmed$file = unlist(lapply(strsplit(trimmed$V1,split="/",fixed=TRUE),function(a) a[6]))
trimmed$sample = unlist(lapply(strsplit(trimmed$file,split="_",fixed=TRUE),function(a) paste(a[1],a[2],sep="_")))
trimmed$read = unlist(lapply(strsplit(trimmed$file,split="_",fixed=TRUE),function(a) a[9]))
trimmed$lane = unlist(lapply(strsplit(trimmed$file,split="_",fixed=TRUE),function(a) a[8]))

bySample = merge(trimmed[trimmed$read=="R1",c("sample","lane","V2","V3")],trimmed[trimmed$read=="R2",c("sample","lane","V2")],by=c("sample","lane"))
bySample$numReadFrag = bySample$V3*2
colnames(bySample) = c("sample","lane","avgFragLength.R1","numReads","avgFragLength.R2","numReadFrag")

rawCounts$lane = unlist(lapply(strsplit(rawCounts$file,split="_",fixed=TRUE),function(a) a[8])) 
rawCounts = rawCounts[grep("_R1",rawCounts$file),]

bySample = merge(rawCounts,bySample,by=c("sample","lane"))
bySample$pctReadsAfterTrim = paste(sprintf("%.1f",round(100*bySample$numReads/bySample$readFrag,1)),"%",sep="")

bySample$numPairedReads = bySample$readFrag
bySample$numReadFragments = bySample$readFrag*2

forPrint2 = bySample[,c("sample","lane","numPairedReads","numReadFragments","avgFragLength.R1","avgFragLength.R2","numReadFrag","pctReadsAfterTrim")]
forPrint2$avgFragLength.R1 = sprintf("%.1f",round(forPrint2$avgFragLength.R1,1))
forPrint2$avgFragLength.R2 = sprintf("%.1f",round(forPrint2$avgFragLength.R2,1))
forPrint2$numReadFrag = prettyNum(forPrint2$numReadFrag,big.mark=",",scientific=FALSE)
forPrint2$numPairedReads = prettyNum(forPrint2$numPairedReads,big.mark=",",scientific=FALSE)
forPrint2$numReadFragments = prettyNum(forPrint2$numReadFragments,big.mark=",",scientific=FALSE)


colnames(forPrint2) = c("sample","lane","Number of Paired-End Reads","Number of Read Fragments","Average Read Fragment Length After Trimming (first read fragment)","Average Read Fragment Length After Trimming (second read fragment)","Number of Read Fragments After Trimming","Percent of Read Fragments That Remained After Trimming")
```

Trimmed Reads/Read Fragments
---------------------------

```{r, results='asis',echo=FALSE}
kable(forPrint2,align=rep("c",ncol(forPrint2)))
```

Total Number of Paired End Reads After Trimming: `r prettyNum(sum(as.numeric(bySample$numReads)),big.mark=",",scientific=FALSE)`  
Total Number of Read Fragments After Trimming:  `r prettyNum(sum(bySample$numReadFrag),big.mark=",",scientific=FALSE)`  
Average Number of Paired End Reads Per Sample After Trimming: `r prettyNum(2*mean(bySample$numReads),big.mark=",",scientific=FALSE)`  

5. Align to ribosomal RNA - RUNNING
---------------------------
```
qsub -q smp /storage/saba/Rat.RI.Brain.totalRNA.batch2/programs/alignTo.rRNA.brain.totalRNA.batch2.sh
```


6.  Convert unmapped.bam file into fastq files - DONE
-----------------------------------------------
```
qsub -q smp /storage/saba/Rat.RI.Brain.totalRNA.batch2/programs/convertToFastQ.batch2.sh
```

